\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-12 k-planar.tex
Pfffft

  Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.  Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.  Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.  Therefore $N_i'$ has fewer inner faces than $N'$.  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.  The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in the closure of $F_i$.
  % The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
  We can now define our desired partition
  \[
      \mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
  \]
  It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.  All that remains is to verify that $\mathcal{P}$ satisfies conditions~(i)--(iv).
  \begin{compactenum}[(i)]
    \item To see that $H$ is planar, observe that $G/\mathcal{P}=G'/\mathcal{P}$\todo{Explain why $G/\mathcal{P}=G'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.  Since $G'$ is planar and planarity is preserved under edge contractions, this implies that $G'/\mathcal{P}$ is planar.

@@

We remark that our proof of \cref{1-planar} generalizes to the class of $d$-framed graphs \cite{BDGGMR}.  An embedded multigraph $G$ is a \defin{$d$-framed graph} if it contains a plane subgraph $G'$ (the \defin{frame}) whose edges are not involved in any crossings and such that each pair of crossing edges in $G$ are contained in a face of $G'$ that has at most $d$ vertices on its boundary.  It is known that the class of $4$-framed multigraphs is exactly the class of $1$-plane multigraphs \cite{BDGGMR}.  Indeed, the graph $G'$ in \cref{induction} is the frame of $G$.  The proof of \cref{induction} can be repeated for any edge-maximal $d$-framed graph $G$ having frame $G'$. The proof in this case has only two differences:
\begin{compactitem}
   \item For any edge $vw$ of $G$ that is not in $G'$, $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$.
   \item The face $\tau$ that contains the Sperner triangle $v_1,v_2,v_3$ may have up to $d$ vertices on its boundary, so up to $d-3$ of these vertices do not belong to $Q_i$ for any $i\in\{1,2,3\}$.
\end{compactitem}
The only effect this has on the final result is that the layered width becomes $3\lfloor d/2\rfloor + d-3$.  This yields the following result:

\begin{thm}
\label{framed}
   Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
\end{thm}

@@

>>>>

  Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.  Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.  Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.  Therefore $N_i'$ has fewer inner faces than $N'$.  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.  The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in the closure of $F_i$.
  % The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
  We can now define our desired partition
  \[
      \mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
  \]
  It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.  All that remains is to verify that $\mathcal{P}$ satisfies conditions~(i)--(iv).
  \begin{compactenum}[(i)]
    \item To see that $H$ is planar, observe that $N/\mathcal{P}=N'/\mathcal{P}$\todo{Explain why $N/\mathcal{P}=N'/\mathcal{P}$} and that each part $P\in\mathcal{P}$ induces a connected subgraph $G'[P]$ of $G'$.  Since $G'$ is planar and planarity is preserved under edge contractions, this implies that $G'/\mathcal{P}$ is planar.

@@

We remark that our proof of \cref{1-planar} generalizes to the class of $d$-framed graphs \cite{BDGGMR} and even $d$-framed multigraphs.  An embedded multigraph $G$ is a \defin{$d$-framed multigraph} if it contains a plane subgraph $G'$ (the \defin{frame}) whose edges are not involved in any crossings and such that each pair of crossing edges in $G$ are contained in a face of $G'$ that has at most $d$ vertices on its boundary.  It is known that every $3$-connected $1$-planar graph is contained in some $4$-framed graph.  The discussion on the structure of $G$ and $G'$ that precedes \cref{induction} shows that every $1$-planar graph is contained in a $4$-framed multigraph.
The proof of \cref{induction} can be repeated for any edge-maximal $d$-framed multigraphs $G$ having frame $G'$. The proof in this case has only two differences:
\begin{compactitem}
   \item For any edge $vw$ of $G$ that is not in $G'$, $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$.
   \item The face $\tau$ that contains the Sperner triangle $v_1,v_2,v_3$ may have up to $d$ vertices on its boundary, so up to $d-3$ of these vertices do not belong to $Q_i$ for any $i\in\{1,2,3\}$.
\end{compactitem}
The only effect this has on the final result is that the layered width becomes $3\lfloor d/2\rfloor + d-3$.  This yields the following result:

\begin{thm}
\label{framed}
   Every $d$-framed multigraph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
\end{thm}

@@

\note{PM}{I no longer agree about the restructuring. See my email essay for an explanation.}

\note{PM}{Apropos the extra clause.  It's related to the orange todo note (argue that $N'/\mathcal{P}=N/\mathcal{P}$).  This doesn't need to be in the statement of the lemma since, unlike \cref{NewLemma} this one gives an $H$-partition of $N$, not $N'$, but something like it needs to be in the proof.}
